MinCall --- MinION end2end deep learning basecaller

### 2017-07-01 paper/main.tex
Normalized E. Coli

\begin{table}[]
	\caption{Alignment specifications of Ecoli R9 basecalled reads using GraphMap}
	\label{tbl:ecoli_rates}
	\centering
	\begin{tabular}{lcccc}
		\toprule
		{} &  \thead{Match \% \\(median)} &  \thead{Mismatch \% \\(median)} &  \thead{Insertion \% \\(median)} &  \thead{Deletion \% \\(median)} \\
		\midrule
		DeepNano   &                  90.254762 &                      6.452852 &                       \textbf{3.274420} &                     11.829965 \\
		Metrichor  &                  90.560455 &                      5.688105 &                       3.660381 &                      8.328271 \\
		Nanonet    &                  90.607674 &                      5.608912 &                       3.652791 &                      8.299046 \\
		MinCall    &                  \textbf{91.408591} &                     \textbf{ 5.019141} &                       3.477739 &                      \textbf{7.471608 }\\
		\bottomrule
	\end{tabular}
\end{table}

\begin{figure}[]
	\begin{center}
		\includegraphics[width=0.8\textwidth]{./imgs/results/ecoli/kde_cigar.png}
		\caption{KDE plot for distribution of percentage of alimnment operations for ecoli}
		\label{fg:ecoli_kde}
	\end{center}
\end{figure}

@@

\begin{table}[]
	\caption{Consensus specifications of Ecoli R9 basecalled reads}
	\label{tbl:spec_ecoli}
	\centering
	\begin{tabular}{lcccccc}
		\toprule
		{} &  \thead{Total called\\\lbrack bp\rbrack} &  \thead{Correctly called\\\lbrack bp\rbrack} &  \thead{Match\\\%} &  \thead{Snp\\\%} &  \thead{Insertion\\\%} &  \thead{Deletion\\\%} \\
		\midrule
		DeepNano  &                  1510244.0 &                      1493242.0 &          98.8742 &         1.0044 &               0.1214 &              0.9041 \\
		Metrichor &                  1515893.0 &                      1502588.0 &          99.1223 &         0.7464 &               0.1313 &              0.6300 \\
		Nanonet   &                  1414237.0 &                      1385515.0 &          97.9691 &         1.5700 &               0.4609 &              1.5158 \\
		resdeep   &                  1517828.0 &                      1506233.0 &          \textbf{99.2361} &         \textbf{0.6474} &               \textbf{0.1165} &             \textbf{ 0.5510 }\\
		\bottomrule
	\end{tabular}
\end{table}

@@

In this evaluation method, the consensus sequence is not calculated from the pileup, but by de novo genome assembly. For this task, fast and accurate de novo genome assembler \emph{ra}\footnote{\url{https://github.com/rvaser/ra}}~\cite{vaser} was used and obtained consensus sequence is compared to the reference using dnadiff present in the Mumer\footnote{\url{https://github.com/garviz/MUMmer}}.
The length of the reference, consensus sequence, number of contigs and percentages of aligned bases from the reference to the query and vice versa are shown in the table \ref{tbl:assembly}. Average identity summarizes how closely does the assembled sequence match the reference. This is run on full ecoli sequence run for 1D template reads (\textasciitilde 160k reads), for both developed tools, Nanonet and Metrichor. Developed tool has shown a small increase in quality of the assembled sequence over Metrichor by offering longer consensus, higher identity percentage, and overall smaller edit distance\footnote{Calculated using \url{https://github.com/isovic/racon/blob/master/scripts/edcontigs.py}}.

\begin{table}[htb]
	\caption{Assembly and consensus results for ecoli}
	\label{tbl:assembly}
	\centering
\begin{tabular}{lccc}
\toprule
&         Metrichor &           MinCall &          Nanonet \\
\midrule
\thead{Ref. genome size (bp)} &           4639675 &           4639675 &            4639675 \\
\thead{Total bases (bp)}      &           4604806 &           \textbf{4614354} &          4600056 \\
\thead{Contigs [\#]}           &                 1 &                 1 &                1 \\
\thead{Aln. bases ref. (bp)}  &  4639641(100.00\%) &  4639612(100.00\%) &  4639031(99.99\%) \\
\thead{Aln. bases query (bp)} &  4604787(100.00\%) &  4614351(100.00\%) &  4599745(99.99\%) \\
\thead{Avg. Identity}         &             98.76 &             \textbf{99.06} &            98.47 \\
\thead{Edit distance}         &             60418 &             \textbf{46686 }&            74341 \\
\bottomrule
\end{tabular}
\end{table}

@@

\begin{table}[]
	\caption{Alignment specifications of Ecoli R9 basecalled reads using BWA mem}
	\label{tbl:ecoli_rates_bwa}
	\centering
	\begin{tabular}{lcccc}
		\toprule
		{} &  \thead{Match \% \\(median)} &  \thead{Mismatch \% \\(median)} &  \thead{Insertion \% \\(median)} &  \thead{Deletion \% \\(median)} \\
		\midrule
		DeepNano   &                  90.254762 &                      6.452852 &                       3.274420 &                     11.829965 \\
		Metrichor  &                  90.595441 &                      6.869543 &                       2.531646 &                      7.567381 \\
		Nanonet    &                  90.988989 &                      6.674760 &                     \textbf{  2.348552} &                      7.698530 \\
		MinCall    &                 \textbf{ 91.470588} &                      \textbf{5.929204 }&                       2.477283 &                     \textbf{ 6.970362} \\
		\bottomrule
	\end{tabular}

>>>>

\begin{table}[]
	\caption{Alignment specifications of E. Coli R9 basecalled reads using GraphMap}
	\label{tbl:ecoli_rates}
	\centering
	\begin{tabular}{lcccc}
		\toprule
		{} &  \thead{Match \% \\(median)} &  \thead{Mismatch \% \\(median)} &  \thead{Insertion \% \\(median)} &  \thead{Deletion \% \\(median)} \\
		\midrule
		DeepNano   &                  90.254762 &                      6.452852 &                       \textbf{3.274420} &                     11.829965 \\
		Metrichor  &                  90.560455 &                      5.688105 &                       3.660381 &                      8.328271 \\
		Nanonet    &                  90.607674 &                      5.608912 &                       3.652791 &                      8.299046 \\
		MinCall    &                  \textbf{91.408591} &                     \textbf{ 5.019141} &                       3.477739 &                      \textbf{7.471608 }\\
		\bottomrule
	\end{tabular}
\end{table}

\begin{figure}[]
	\begin{center}
		\includegraphics[width=0.8\textwidth]{./imgs/results/ecoli/kde_cigar.png}
		\caption{KDE plot for distribution of percentage of alimnment operations for E. Coli}
		\label{fg:ecoli_kde}
	\end{center}
\end{figure}

@@

\begin{table}[]
	\caption{Consensus specifications of E. Coli R9 basecalled reads}
	\label{tbl:spec_ecoli}
	\centering
	\begin{tabular}{lcccccc}
		\toprule
		{} &  \thead{Total called\\\lbrack bp\rbrack} &  \thead{Correctly called\\\lbrack bp\rbrack} &  \thead{Match\\\%} &  \thead{Snp\\\%} &  \thead{Insertion\\\%} &  \thead{Deletion\\\%} \\
		\midrule
		DeepNano  &                  1510244.0 &                      1493242.0 &          98.8742 &         1.0044 &               0.1214 &              0.9041 \\
		Metrichor &                  1515893.0 &                      1502588.0 &          99.1223 &         0.7464 &               0.1313 &              0.6300 \\
		Nanonet   &                  1414237.0 &                      1385515.0 &          97.9691 &         1.5700 &               0.4609 &              1.5158 \\
		resdeep   &                  1517828.0 &                      1506233.0 &          \textbf{99.2361} &         \textbf{0.6474} &               \textbf{0.1165} &             \textbf{ 0.5510 }\\
		\bottomrule
	\end{tabular}
\end{table}

@@

In this evaluation method, the consensus sequence is not calculated from the pileup, but by de novo genome assembly. For this task, fast and accurate de novo genome assembler \emph{ra}\footnote{\url{https://github.com/rvaser/ra}}~\cite{vaser} was used and obtained consensus sequence is compared to the reference using dnadiff present in the Mumer\footnote{\url{https://github.com/garviz/MUMmer}}.
The length of the reference, consensus sequence, number of contigs and percentages of aligned bases from the reference to the query and vice versa are shown in the table \ref{tbl:assembly}. Average identity summarizes how closely does the assembled sequence match the reference. This is run on full E. Coli sequence run for 1D template reads (\textasciitilde 160k reads), for both developed tools, Nanonet and Metrichor. Developed tool has shown a small increase in quality of the assembled sequence over Metrichor by offering longer consensus, higher identity percentage, and overall smaller edit distance\footnote{Calculated using \url{https://github.com/isovic/racon/blob/master/scripts/edcontigs.py}}.

\begin{table}[htb]
	\caption{Assembly and consensus results for E. Coli}
	\label{tbl:assembly}
	\centering
\begin{tabular}{lccc}
\toprule
&         Metrichor &           MinCall &          Nanonet \\
\midrule
\thead{Ref. genome size (bp)} &           4639675 &           4639675 &            4639675 \\
\thead{Total bases (bp)}      &           4604806 &           \textbf{4614354} &          4600056 \\
\thead{Contigs [\#]}           &                 1 &                 1 &                1 \\
\thead{Aln. bases ref. (bp)}  &  4639641(100.00\%) &  4639612(100.00\%) &  4639031(99.99\%) \\
\thead{Aln. bases query (bp)} &  4604787(100.00\%) &  4614351(100.00\%) &  4599745(99.99\%) \\
\thead{Avg. Identity}         &             98.76 &             \textbf{99.06} &            98.47 \\
\thead{Edit distance}         &             60418 &             \textbf{46686 }&            74341 \\
\bottomrule
\end{tabular}
\end{table}

@@

\begin{table}[]
	\caption{Alignment specifications of E. Coli R9 basecalled reads using BWA mem}
	\label{tbl:ecoli_rates_bwa}
	\centering
	\begin{tabular}{lcccc}
		\toprule
		{} &  \thead{Match \% \\(median)} &  \thead{Mismatch \% \\(median)} &  \thead{Insertion \% \\(median)} &  \thead{Deletion \% \\(median)} \\
		\midrule
		DeepNano   &                  90.254762 &                      6.452852 &                       3.274420 &                     11.829965 \\
		Metrichor  &                  90.595441 &                      6.869543 &                       2.531646 &                      7.567381 \\
		Nanonet    &                  90.988989 &                      6.674760 &                     \textbf{  2.348552} &                      7.698530 \\
		MinCall    &                 \textbf{ 91.470588} &                      \textbf{5.929204 }&                       2.477283 &                     \textbf{ 6.970362} \\
		\bottomrule
	\end{tabular}
